---
title: "Clean Allocated Values / Combine the datasets  (cl_01)"
date: "2025-06-14"
format: html
execute:
  warning: false
  message: false
  eval: false
---

## 🎯 Purpose

For this file, your goal is:
1. Remove allocated values that are not meaningful for further analysis
2. Merge or split datasets as needed to prepare for the next stage of cleaning and transformation

## Clean up all allocated values

```{r}

# Load data
data_2000 <- read.xlsx("data/outcome/2000.dta")

# Define quality variable list
qvarlist <- c("var_1", "var_2"...)

# Remove observations where quality flags are greater than 3 (this is your to-go), then drop the quality variables
# Note: Definitions for allocated values may vary by variable, so greater than 3 may not be universal, so always double-check the documentation.

for(qvar in qvarlist) {
  if(qvar %in% names(data_2000)) {
    data_2000 <- data_2000 %>%
      filter(!!sym(qvar) <= 3) %>%
      select(-!!sym(qvar))
  }
}

# Save cleaned data
write.xlsx(data_2000, "data/outcome/2000_i.xlsx")
```

## Merging the values

Most of the cases, if you not download your dataset from the customized dataset, you will have a zip of documents, and you may need to manually merge them. 

For example, in the [BLS Quarterly Census of Employment and Wages](https://www.bls.gov/cew/), when you download, you may have series of datasets by year, so right now

```{r}
years <- c(2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022)
```


To merge the files together in R,there are two ways for your to do this:
1. you can use the bind_rows function, but this approach is only convenient when you don't have so much of the files, and the process can be a little bit of messy

```{r}
qcew_data_2009 <- read_csv("data/outcome/2009_qcew.csv")
qcew_data_2010 <- read_csv("data/outcome/2010_qcew.csv")
...

qcew_combined <- bind_rows(qcew_data_2009, qcew_data_2010, ...)

```

2. When you have many of the years, you can setup a list, and directly merge all the files in the list

```{r}
file_list <- paste0("data/outcome/", years, "_qcew.csv")

qcew_combined <- map_dfr(file_list, read_csv)
```

📌 If you encounter an error during this step, check:
(1) whether the file names are correct, and
(2) whether the files have matching column structures.

💭 In this section, I demonstrated how to combine separate files into a single dataset. Now, try the opposite—splitting a dataset into parts.
Also, take a moment to reflect:

 Under what assumptions is it appropriate to combine datasets?
 
 When might it be better to keep them separate?