---
title: "Data Extraction and Renaming (cl_00)"
date: "2025-06-14"
format: 
  html:
    toc: true
    code-fold: false
    theme: cosmo
  pdf:
    toc: true
execute:
  warning: false
  message: false
  eval: false
---

## 🎯 Purpose

In this step, we load the raw dataset, convert it into a more flexible format, and rename it to make it easier to reference in later steps. Think of this as preparing your workspace before you start cleaning.

## Setup

We'll start by loading the required libraries for reading Stata, Excel, and csv files:

```{r setup}
#| label: setup
#| include: true

# Load required libraries
library(haven)      # for reading/writing Stata files
library(openxlsx)   # for reading/writing xlsx files
library(readr)      # for reading/writing csv files
library(data.table) # good for large csv files
```

### Load the data and rename it

Most datasets are available as .csv files, which we can use directly in R.

📌 While R supports both .csv and .xlsx formats, .csv is generally preferred because it's lightweight, universally compatible, and easier to integrate with other tools and platforms[^1].

```{r}
#| label: load-source-data

# Load the source IPUMS data file; for example, if I am working with data in 2000, I will directly name the file as 2000 or 2000_acs
data_2000 <- read_csv("data/source/your_file_name_here.csv")


write_csv(data_2000, "data/outcome/2000.csv")

```

If the files are large, you can also choose to use fread

```{r}
data_2000 <- fread("data/source/your_file_name_here.csv")


fwrite(data_2000, "data/outcome/2000.csv")
```

If you have a list of files (I use BLS data as an example here), you can 

```{r}
years <- c(2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022)

# Loop through each year
for(year in years) {
  
  # Define file paths
  input_file <- paste0("data/source/QCEW/fast_food sector/", year, "_Limited-service restaurants.csv")  #you will be list the name as the way they downloaded on your local
  output_file <- paste0("data/outcome/", year, "_bls.csv")
  
  # Read CSV file
  data <- read_csv(input_file)
  
  # Save as CSV file with new name
  write_csv(data, output_file)
}
```

 [^1]: For a comparison of `.csv` vs `.xlsx` in data workflows, see this LinkedIn article: [CSV vs Excel: Pros and Cons](https://www.linkedin.com/advice/0/what-pros-cons-using-csv-vs-excel-data-validation-analysis)

