---
title: "Data Organization"
date: "2025-06-14"
format: html
execute:
  warning: false
  message: false
  eval: false
---

## 📋 Overview of the Data Cleaning Procedure: the \_master file

The \_master file serves as the central script that organizes all your data cleaning and analysis steps. In most research projects, you'll eventually need to submit your code---having a well-structured master file ensures that reviewers (or your future self) can reproduce the results easily by running just one file.

## Setup

The Functions of the basic packages

```{r setup}
#| label: setup
#| include: true

# Data manipulation and transformation
library(dplyr)      # Grammar of data manipulation - filter, select, mutate, summarize, group_by

# Data import/export
library(readr)      # Fast and user-friendly reading of rectangular data (CSV, TSV, etc.)
library(haven)      # Reading and writing data from statistical software (Stata, SPSS, SAS)

# Data reshaping and tidying
library(tidyr)      # Tools for reshaping data - pivot_longer, pivot_wider, separate, unite

# Data visualization
library(ggplot2)    # Grammar of graphics for creating complex, layered plots

# Reproducible reporting
library(knitr)      # Dynamic report generation - combine R code with text
library(here)       # Portable file path construction relative to project root
library(quarto)     # Scientific publishing system for documents, presentations, websites

# Display the result of analysis 
library(stragazer)  # Display the regression result clealy in latex format
```

### Set working directory

We typically define the working directory in the master file, so it only needs to be set once. This simplifies code submission---reviewers can run everything without adjusting multiple scripts.

```{r}
# change this to your local directory
setwd("/Users/zhaoxiaoxiao/Desktop/2000 Educ_Return_R")
```

## Part I: Data Cleaning Process

### Step 1: Data Import and Renaming {#sec-step1}

Rename the dataset using a clear, descriptive name, and convert it to the desired format at this stage.

```{r step1}
#| label: data-import
#| eval: false

# Execute data import and renaming script
quarto_render("program/cl_00.qmd")
```

### Step 2: Clean Allocated Values / Combine the datasets {#sec-step2}

In this step, we remove any allocated or imputed values to maintain data quality and transparency.

If the dataset does not contain allocated values, this step is used to merge multiple raw files---for example, when data comes in separate parts by geography or time.

📌 Be sure to clearly document the merging logic and assumptions, especially if the source files were not designed to be combined automatically.

```{r step2}
#| label: clean-allocated
#| eval: false

# Clean up all allocated values
quarto_render("program/cl_01.qmd")
```

### Step 3: Variable Generation and Transformation {#sec-step3}

This is where most of the data cleaning happens---you'll recode variables, create new ones, and keep only what you need for analysis.

```{r step3}
#| label: variable-generation
#| eval: false

# Generate and keep relevant variables
quarto_render("program/cl_02.qmd")
```

**Key transformations include:**

a.  **Missing Value Treatment**: Replace missing value codes into na
b.  **Top-coding Indicator**: Generate variable indicating whether the value is topcoded (for example, income)
c.  **Inflation Adjustment**: Adjust for inflation using appropriate price indices (we ususally need this step when we talk about price and income)
d.  **Recode and Create Variables**: Convert categorical values to numeric, group variables, or build new ones for analysis.
e.  **Save What You Need**: Keep only the cleaned, final set of variables for modeling or visualization

### Step 4: Sample Restrictions {#sec-step4}

Apply sample restrictions to create the final analytical dataset.

```{r step4}
#| label: sample-restrictions
#| eval: false

# Apply sample restrictions
quarto_render("program/cl_03.qmd")
```

**Purpose:** Implement inclusion/exclusion criteria to define the population of interest.

## Data Analyzing

```{r}
# Run your most basic analysis
quarto_render("program/an_01.qmd")
```

## Data Processing Summary

The data cleaning process follows a systematic approach:

1.  **Raw Data Import** → Standardized file structure
2.  **Quality Control** → Remove problematic allocated values\
3.  **Feature Engineering** → Create analysis variables
4.  **Sample Definition** → Apply research-specific restrictions

## File Structure

```         
project/
├── program/
|   ├── _master.qmd  # Organize all the files under, set the main working directory
│   ├── cl_00.qmd    # Data import and renaming
│   ├── cl_01.qmd    # Clean allocated values
│   ├── cl_02.qmd    # Variable generation
│   ├── cl_03.qmd    # Sample restrictions
|   ├── an_01.qmd    # This will be your basic model result
|   └── ...          # Other model that you want to build
├── data/
|   ├── source     # storage of the raw datasets
|   └── output     # storage of cleaned datasets
├── log/           # storage of the results of data analysis 
├── graph/         # storage of graphs
└── reference/     # reference of your research
```

## 📌 Note

1.  If you're working with more than one raw dataset, it's good practice to create a separate cleaning script for each one. For example:

```{r}
quarto_render("program/cl_01_acs.qmd")  # Load the acs data

quarto_render("program/cl_01_cps.qmd")  # Load the cps data
```

If the datasets need to be merged, write a dedicated merge script

```{r}
quarto_render("program/cl_04_merge.qmd")
```

2.  Your \_master script should do more than just run your code---it should also:

<!-- -->

(i) Describe the purpose of each cleaning and analysis file
(ii) Clearly label each stage of the workflow (e.g., cleaning, merging, modeling)
(iii) Help collaborators or reviewers understand the flow of your project at a glance

<!-- -->

