---
title: "example_1"
format: html
execute:
  warning: false
  message: false
  eval: false
---

## 🔍 Example 1: Replicating AK Paper

For this example, we’ll walk through a simplified replication exercise based on the paper [“Does Compulsory School Attendance Affect Schooling and Earnings?” by Angrist and Krueger (1991)](http://piketty.pse.ens.fr/files/AngristKrueger1991.pdf). For this exercise, we’ll clean a dataset from the ACS to prepare it for studying the relationship between years of education and income. Since this paper used the birth quarter as the instrument as the years of the education, we will go through the practice of cleaning out year of birth and quarter of birth as well.  

### Data Import (cl_00)

For this task, we’ll download data from IPUMS and split it by year. The purpose of this approach is to allow us to test the model’s significance separately for each individual year later on.

```{r}
# usa_00012 is the file name when you directly download from IPUMS

data_acs <- read_csv("data/source/usa_00012.csv")

#Make sure the 'year' column is treated as a factor or numeric
data_acs$year <- as.integer(data_acs$year)

# Split the data into a list of datasets, one per year
# Now data_by_year is a named list, e.g. data_by_year[["2001"]], data_by_year[["2002"]], etc.

data_by_year <- split(data_acs, data_acs$year)

for (yr in names(data_by_year)) {
  write.csv(data_by_year[[yr]], paste0("data/outcome/", yr, ".csv"), row.names = FALSE)
}  
  
```

### Clean out the Allocated Values (cl_01)

In this example, we need to work with 1980 data, which uses different set of allocated values compared to other years. Because of this, we handle the cleaning process for 1940 separately. This is also one of the main reasons we split the dataset by year in the first place— the more years of data we include, the more likely it is that coding differences or structural changes will occur across time.

```{r}
# Load 1980 data
data_1980 <- read_csv("data/outcome/1980.csv")

#define the allocated values 
qvarlist <- c("qage", "qmarst", "qsex", "qbpl", "qrace", "qeduc", 
              "qschool", "qclasswk", "qempstat", "quhrswor", 
              "qwkswork1", "qincwage")


# Process each variable in the list
for (i in qvarlist) {
  # Drop observations where the variable value is greater than 3
  # (equivalent to 'drop if `i'>3')
  data <- data[data[[i]] <= 3 | is.na(data[[i]]), ]
  
  # Drop the variable from the dataset
  # (equivalent to 'drop `i'')
  data <- data[, !names(data) %in% i]
}


# Save cleaned 1980 data
write_csv(data_1980, "data/outcome/1980_i.csv")
```


```{r}

years <- c(2005, 2006, 2007, 2019, 2020, 2021, 2023)

qvarlist <- c("qage", "qmarst", "qsex", "qbpl", "qrace", "qeduc", 
              "qschool", "qclasswk", "qempstat", "quhrswor", 
              "qwkswork1", "qincwage")

# Loop through each year (equivalent to 'foreach k in `r(numlist)'')
for (k in years) {
  
  # Load the data for current year (equivalent to 'use data/outcome/`k'.dta,clear')
  data <- read_dta(paste0("data/outcome/", k, ".dta"))
  
  # Process each variable in the list
  for (i in qvarlist) {
    # Drop observations where the variable value is greater than 3
    # (equivalent to 'drop if `i'>3')
    data <- data[data[[i]] <= 3 | is.na(data[[i]]), ]
    
    # Drop the variable from the dataset
    # (equivalent to 'drop `i'')
    data <- data[, !names(data) %in% i]
  }
  
  # Save the processed data (equivalent to 'save data/outcome/`k'_i.dta,replace')
  write_csv(data, paste0("data/outcome/", k, "_i.dta"))
}
```

### Data Cleaning (cl_02)

For this dataset, our dependent variables will include: annual wage, weekly wage, and weeks worked in the previous year.

The key independent variables are: years of education, birth year, and birth quarter.

Additional demographic controls include: race, marital status, employment status, age group, working experience, and gender.

#### Replacing code for NA values with na

```{r}
years <- c(1980, 2005, 2006, 2007, 2019, 2020, 2021, 2023)

# Loop through each year (equivalent to 'foreach i in `r(numlist)'')
for (i in years) {
  
  # Load the processed data (equivalent to 'use data/outcome/`i'_i,clear')
  data <- read_csv(paste0("data/outcome/", i, "_i.csv"))
  
  # Replace incwage with missing value if >= 999999
  # (equivalent to 'replace incwage = . if incwage >= 999999')
  data$incwage[data$incwage >= 999999] <- NA
  
  # Save the data (equivalent to 'save data/outcome/`i'_i,replace')
  write_csv(data, paste0("data/outcome/", i, "_i.dta"))
  
  # Print progress (optional)
  cat("Processed incwage for year:", i, "\n")
}
```

#### Generate a variable indicating whether or not the Individual's wage is top-coded

For this variable, each year can have different definition for the topcode. 
```{r}
data_1980 <- read_csv("data/outcome/1980_i.csv") %>%
  mutate(topcoded = as.numeric(incwage >= 75000 & incwage < 999999 & !is.na(incwage)))

# For 2005  
data_2005 <- read_csv("data/outcome/2005_i.csv") %>%
  mutate(topcoded = as.numeric(incwage >= 175000 & incwage < 999999 & !is.na(incwage)))
...
```

#### Adjust the annual wage by inflation

```{r}
data_1980 <- read_csv("data/outcome/1980_i.csv")

data_1980$incwage <- data_1980$incwage * 2.314

write_csv(data_1980, "data/outcome/1980_i.csv")

data_2005 <- read.csv("data/outcome/2005_i.csv")

data_2005$incwage <- data_2005$incwage * 0.853

write.csv(data_2005, "data/outcome/2005_i.csv", row.names = FALSE)

...
```

#### Recode the Education variable and generate working experience

As mentioned earlier in the data access section, maintaining consistency of variables across years is important. This is a good example: the variable for years of education completed isn’t directly available in the ACS. Instead, we can use two closely related variables as proxies:
[highgrade](https://usa.ipums.org/usa-action/variables/HIGRADE#codes_section)(the highest grade of schooling, easy to clean for years of education, but not available after 1980) and [educ](https://usa.ipums.org/usa-action/variables/EDUC#description_section)(the education attainment, it is available for more years, but harder to clean). 

```{r}
data_1980 <- read_csv("data/outcome/1980_i.csv")

  data_1980$higrade[data_1980$higrade == 0] <- NA
  data_1980$higrade[data_1980$higrade %in% c(1, 2, 3)] <- 0

# Generate schooling variable (equivalent to 'g schooling = higrade')
  data_1980$schooling <- data_1980$higrade

# Adjust schooling (equivalent to 'replace schooling = schooling - 3 if schooling>3')
  data_1980$schooling[data_1980$schooling > 3 & !is.na(data_1980$schooling)] <- 
  data_1980$schooling[data_1980$schooling > 3 & !is.na(data_1980$schooling)] - 3

# Generate experience (equivalent to 'g exp = age - schooling - 6')
  data_1980$exp <- data_1980$age - data_1980$schooling - 6

# Replace negative experience with 0 
  data_1980$exp[data_1980$exp < 0 & !is.na(data_1980$exp)] <- 0

# Generate experience squared 
  data_1980$exp2 <- data_1980$exp * data_1980$exp

# Generate age squared 
  data_1980$age2 <- data_1980$age * data_1980$age

# Save the data (equivalent to 'save data/outcome/1980_i,replace')
  write_csv(data_1980, "data/outcome/1980_i.csv")
```


```{r}
years <- c(2005, 2006, 2007, 2019, 2020, 2021, 2023)

for (i in years) {
  
  # Load the data 
  data <- read_dta(paste0("data/outcome/", i, "_i.dta"))
  
  # Generate schooling from educd 
  data$schooling <- data$educd
  
  # Recode schooling values 
  schooling_map <- c(
  "0" = NA, "1" = NA, "2" = NA,
  "10" = 2.5, "21" = 5.5, "24" = 7.5, "30" = 9,
  "40" = 10, "50" = 11, "61" = 12, "62" = 12, "63" = 12, "64" = 12,
  "65" = 13, "71" = 13, "81" = 14, "101" = 16, "114" = 18, "115" = 16, "116" = 16
)
  
  data$schooling <- schooling_map[as.character(data$educd)]
  
  data$exp <- data$age - data$schooling - 6
  
  data$exp[data$exp < 0 & !is.na(data$exp)] <- 0
  
  data$exp2 <- data$exp * data$exp
  
  data$age2 <- data$age * data$age
  
  write_csv(data, paste0("data/outcome/", i, "_i.csv"))
  
}
```

#### Recode the weeks usually worked, weekly earnings, and hourly earning

```{r}

years <- c(1980, 2005, 2006, 2007, 2019, 2020, 2021, 2023)

# ****************************************************
# Generate weekly wage
# ****************************************************

for (i in years) {
  
  # Load the data 
  data <- read_csv(paste0("data/outcome/", i, "_i.csv"))
  
  # Recode wkswork1 
  data$wkswork1[data$wkswork1 == 0] <- NA
  
  # Generate weeks variable 
  data$weeks <- data$wkswork1
  
  # Generate weekly wage 
  data$wagewk <- data$incwage / data$weeks
  
  # Save the data 
  write_csv(data, paste0("data/outcome/", i, "_i.csv"))
  
}

# ****************************************************
# Generate hourly wage
# ****************************************************

for (i in years) {
  
  # Load the data 
  data <- read_csv(paste0("data/outcome/", i, "_i.csv"))
  
  # Recode uhrswork 
  data$uhrswork[data$uhrswork == 0] <- NA
  
  # Generate topcoded hours indicator 
  data$topcoded_uhr <- as.numeric(data$uhrswork == 99 & !is.na(data$uhrswork))
  
  # Generate hourly wage 
  data$wagehr <- data$incwage / (data$weeks * data$uhrswork)
  
  # Save the data 
  write_csv(data, paste0("data/outcome/", i, "_i.csv"))
}
```

💭 Exercise: Can you write the code to generate each individual's year of birth and quarter of birth based on the available variables?

#### Generate Other Demographics

```{r}
years <- c(1980, 2005, 2006, 2007, 2019, 2020, 2021, 2023)

for (i in years) {
  
data <- read_dta(paste0("data/outcome/", i, "_i.dta"))
  
# ****************************************************
# Marital Status
# ****************************************************
  
    # Generate married indicator 
    data$marr <- as.numeric(data$marst == 1 | data$marst == 2)
  
  
# ****************************************************
# Have a child under age 5
# ****************************************************
  
   # Generate child under 5 indicator 
  data$child5 <- as.numeric(data$nchlt5 > 0 & !is.na(data$nchlt5))
  
  ...
  
# ****************************************************
# Keep the relevant variables
# ****************************************************
  
  # Define variable lists 
  indvar <- c("schooling", "exp", "exp2", "topcoded")
  
  depvar <- c("incwage", "wagewk", "wagehr")
  
  othervar <- c("year", "sample", "serial", "pernum",
                "gq", "statefip", "age", "ageg", "age2",
                "marr", "child5", ...)
  
  pweight <- c("perwt")
  
  # Combine all variables to keep
  vars_to_keep <- c(depvar, indvar, othervar, pweight)
  
  # Method 1: Shortest - direct subsetting with error handling
  data_subset <- data[, intersect(vars_to_keep, names(data)), drop = FALSE]
  
  # Save the data (equivalent to 'save data/outcome/`i'_pre,replace')
  write_csv(data_subset, paste0("data/outcome/", i, "_pre.csv"))

}
  
```

💭 The code above shows how to create a few demographic variables. Now, try extending it by generating the remaining ones—think about how you would code the rest based on the available data.

### Sample restrictions

```{r}
years <- c(1980, 2005, 2006, 2007, 2019, 2020, 2021, 2023)

for (i in years) {
  
  # Load the data 
  data <- read_dta(paste0("data/outcome/", i, "_pre.dta"))
  
  # if we limit the gender to male
  data <- data[data$male == 1 & !is.na(data$male), ]
  
  # keep only the individual with positive income
  data <- data[data$incwage > 0 & !is.na(data$incwage), ]
  
  #touse the missing variables
    data <- data %>%
    mutate(
      touse = complete.cases(select(., incwage, schooling, exp, exp2,
                                   topcoded, perwt, ...))
    ) %>%
    filter(touse)
    
  # If we want to have a unlinear model, you can post the log tranformation here 

  # Generate log wage 
  data$lnwage <- log(data$incwage)
    
  write_dta(data, paste0("data/outcome/", i, "_final.dta"))
  
}



```
