---
title: "Data Organization"
author: "Xiaoyi"
date: "2009-05-21"
format: html
execute:
  warning: false
  message: false
  eval: false
---

## Overview of the Data Cleaning Procedure

## Setup: _master

The Functions of the basic packages

```{r setup}
#| label: setup
#| include: true

# Data manipulation and transformation
library(dplyr)      # Grammar of data manipulation - filter, select, mutate, summarize, group_by

# Data import/export
library(readr)      # Fast and user-friendly reading of rectangular data (CSV, TSV, etc.)
library(haven)      # Reading and writing data from statistical software (Stata, SPSS, SAS)

# Data reshaping and tidying
library(tidyr)      # Tools for reshaping data - pivot_longer, pivot_wider, separate, unite

# Data visualization
library(ggplot2)    # Grammar of graphics for creating complex, layered plots

# Reproducible reporting
library(knitr)      # Dynamic report generation - combine R code with text
library(here)       # Portable file path construction relative to project root
library(quarto)     # Scientific publishing system for documents, presentations, websites
```

### Set working directory

We usually set this up in the master file, so the working directory only needs to be changed once. This makes it easier when submitting your code---reviewers can run everything without modifying multiple files.

```{r}
# change this to your local directory
setwd("/Users/zhaoxiaoxiao/Desktop/2000 Educ_Return_R")
```

## Part I: Data Cleaning Process

### Step 1: Data Import and Renaming {#sec-step1}

Download data from IPUMS in Stata format and rename by year.

```{r step1}
#| label: data-import
#| eval: false

# Execute data import and renaming script
quarto_render("program/cl_00.qmd")
```

**Purpose:** Import raw data files and establish consistent naming conventions by survey year.

### Step 2: Clean Allocated Values {#sec-step2}

Remove or flag all allocated values in the dataset.

```{r step2}
#| label: clean-allocated
#| eval: false

# Clean up all allocated values
quarto_render("program/cl_01.qmd")
```

**Purpose:** Ensure data quality by handling allocated/imputed values that may introduce bias in our analysis.

### Step 3: Variable Generation and Transformation {#sec-step3}

This step involves multiple sub-processes to create analysis-ready variables:

```{r step3}
#| label: variable-generation
#| eval: false

# Generate and keep relevant variables
quarto_render("program/cl_02.qmd")
```

**Key transformations include:**

a.  **Missing Value Treatment**: Replace N/A values with proper missing value codes
b.  **Top-coding Indicator**: Generate variable indicating whether individual's wage is top-coded
c.  **Inflation Adjustment**: Adjust wages for inflation using appropriate price indices
d.  **Education Variables**:
    -   Recode education variable
    -   Generate potential experience
    -   Generate education variable for spouse
e.  **Earnings Variables**: Create standardized earnings measures
f.  **Covariates**: Generate and retain other relevant control variables

### Step 4: Sample Restrictions {#sec-step4}

Apply sample restrictions to create the final analytical dataset.

```{r step4}
#| label: sample-restrictions
#| eval: false

# Apply sample restrictions
quarto_render("program/cl_03.qmd")
```

**Purpose:** Implement inclusion/exclusion criteria to define the population of interest for the returns to education analysis.

## Data Processing Summary

The data cleaning process follows a systematic approach:

1.  **Raw Data Import** → Standardized file structure
2.  **Quality Control** → Remove problematic allocated values\
3.  **Feature Engineering** → Create analysis variables
4.  **Sample Definition** → Apply research-specific restrictions


## File Structure

```         
project/
├── program/
|   ├── _master.R  # Organize all the files under, set the main working directory
│   ├── cl_00.R    # Data import and renaming
│   ├── cl_01.R    # Clean allocated values
│   ├── cl_02.R    # Variable generation
│   └── cl_03.R    # Sample restrictions
├── data/
|   ├── source     # storage of the raw datasets
|   └── output     # storage of cleaned datasets
├── log/           # storage of the results of data analysis 
├── graph/         # storage of graphs
└── reference/     # reference of your research
```
